What paradigms can webcam eye-tracking be used for? Attempted replications of 5 ``classic'' cognitive science experiments

### 2023-08-24 group-d/E4-analysis-replication-in-lab.Rmd
updated intro

```{r , include=FALSE}
library(papaja)
library(jsonlite)
library(purrr)
#library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
```

@@

```{r E4-in-lab-load-data}
user.data.files <- list.files('data/in-lab-merced', pattern = "pushkin_users_*",full.names = TRUE)

all.user.data <- map_dfr(user.data.files, readRDS)

@@

#data.tables = readRDS("data/in-lab-merced/lex_stimulusResponses_2023-05-01.rds")
all.data <- map_dfr(data.files, readRDS) %>% 
  bind_rows(readRDS("data/in-lab-merced/verb_bias_ET_data_10_13_2022.rds")) %>% 
  arrange(user_id, trial_index, time_elapsed) %>% 
  group_by(user_id, trial_index) %>% 
  mutate(subjnum = 1:n(),
         unique_id = str_c(subjnum,"__", user_id)) %>% 
  ungroup() %>% 
  arrange(unique_id, trial_index)

@@

*TO DO:* cross ref with angela's list of participant times/dates

>>>>

```{r , include=FALSE}
# THIS IS A VERY MESSY RMD IN WHICH I TRIED TO ANALYSE THE IN-LAB REPLICATION DATA: DO NOT USE, I WILL EVENTUALLY REMOVE IT - RR
# subject identifiers did not get saved (and some other important stuff too) so the data cannot be analyzed in a reasonable way
library(papaja)
library(jsonlite)
library(purrr)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lmerTest)
library(afex)
library(forcats)
library(broom.mixed)
library(lubridate)
```

@@

```{r E4-in-lab-load-data-first-batch}

#  without dates for user ids
add_order = function(trial_index, lag_trial_index, lag_trial_order, verbose = F){

  if( length(lag_trial_index)==0 | length(lag_trial_order)==0 ){
    trial_order = 0
  }else if(trial_index == lag_trial_index){
    trial_order = lag_trial_order
  }else if(trial_index < lag_trial_index){
    trial_order = 0
  }else if(trial_index > lag_trial_index){
    trial_order = lag_trial_order+1
  }

  if (verbose){
    cat("\n trial index: ",trial_index,"\n ")
    cat("lag trial index: ",lag_trial_index,"\n ")
    cat("lag trial order: ",lag_trial_order,"\n ")
    cat("new trial order: ", trial_order)
  }
   return(trial_order)
}
#add_order(3,1,1,T)

batch1<-readRDS("data/in-lab-merced/verb_bias_ET_data_10_13_2022.rds") %>% 
  tibble::rownames_to_column("row_order") %>% 
  mutate(row_order = as.numeric(row_order)) %>% 
  separate(internal_node_id, into=c('node0','node1', 'node2', 'node3'), sep = "-",remove = F) %>%
  mutate(across(starts_with("node"),as.numeric)) %>% 
  arrange(user_id, row_order) #%>% 
  #group_by(user_id, node0,node1,node2,node3) %>% 
 # group_by(user_id) %>%
 # mutate(trial_order  = 0)

subj_orders = vector()
subjnum = 1
for (i in 1:length(batch1$row_order)){
  if(str_detect(batch1$stimulus[i], "hypothesis")){
    subjnum = subjnum+1
  }
  subj_orders[i] = subjnum
}
#add_order(3,1,1,T)

batch1z = batch1 
batch1z$subj_order = subj_orders

batch1z = batch1 %>% 
     arrange(user_id, trial_index, row_order) %>% 
    # mutate(prev_trial_end = time_elapsed - rt - 2) %>% 
     group_by(user_id, trial_index) %>% 
     mutate(daily_subjnum = 1:n(),
             subj_order = str_c(daily_subjnum,"__", user_id)) #%>%

dayxz = batch1z %>% filter(user_id == 'c402ba0c-7a9d-4258-addd-149223724707')  %>% 
  select(-starts_with('webgazer')) %>% 
  ungroup() %>% 
  mutate(time_lag = as.numeric(time_elapsed) - as.numeric(lag(time_elapsed)))

```
# BATCH1Z
```{r}
batch1z.task.data = batch1z %>%
  dplyr::filter(compatibility != 'NA', compatibility != 'filler') %>%
  dplyr::select(subj_order, trial_index, rt, images, webgazer_data, mouse_events, compatibility, audio, target_instrument, target_animal, webgazer_targets)

# Add a column that uniquely identifies the combination of images and audio shown on the screen
batch1z.trialID.data <- batch1z.task.data %>%
  group_by(audio, images) %>%
  slice(1) %>%
  select(audio, images) %>%
  ungroup() %>%
  mutate(trialID = 1:n())

batch1z.task.data <- batch1z.task.data %>%
  left_join(batch1z.trialID.data)

batch1z.eyetracking.data <- batch1z.task.data %>%
  tidyr::unpack(webgazer_targets) %>%
  tidyr::unpack(c(`#jspsych-free-sort-draggable-0`, `#jspsych-free-sort-draggable-1`, `#jspsych-free-sort-draggable-2`, `#jspsych-free-sort-draggable-3`), names_sep=".") %>%
  unnest_wider(webgazer_data) %>% 
  unnest_longer(col = c(x, y, t))

batch1z.mousetracking.data <- batch1z.task.data %>%
  unnest(mouse_events)
```
```{r E4-categorizing-first-mouse-moves}
#First, create a data set with each trial, adding column for which object the mouse moves over first.
# WTF IS HAPPENING with TRIAL ID?? 
batch1z.first.move <- batch1z.mousetracking.data %>%
  select(subj_order, trialID, object, type, compatibility, target_instrument, target_animal, images, audio) %>%
  group_by(subj_order, trialID) %>%
  filter(type=="enter") %>%
  slice(1) %>%
  rowwise() %>%
  mutate(which_loc = which(images == object)-1) %>%
  ungroup() %>%
  mutate(instrument_loc = str_sub(target_instrument, start=-1, end=-1)) %>%
  mutate(animal_loc = str_sub(target_animal, start=-1, end=-1)) %>%
  mutate(first.move.type = case_when(
    which_loc == instrument_loc ~ 'instrument',
    which_loc == animal_loc ~ 'animal',
    TRUE ~ 'other')) %>%
  mutate(is.mouse.instrument = if_else(first.move.type == 'instrument', 1, 0)) %>%
  mutate(compatibility = factor(compatibility))
```

```{r E4-summarizing-first-mouse-moves-for-plotting}
# Summarize the data by subject, calculating proportion of trials the the first move was to the animal, instrument, or other.
batch1z.first.move.subject.summary <- batch1z.first.move %>%
  group_by(subj_order, compatibility) %>%
  summarize(prop.animal = mean(first.move.type == 'animal'),
            prop.instrument = mean(first.move.type == 'instrument'),
            prop.other = mean(first.move.type == 'other'),
            n_trials= n()) %>%
  filter(n_trials > 9) %>% ## CAHNGE WHEN THERE'S A BETTER WAY TO CROSSREF
  pivot_longer(c('prop.animal', 'prop.instrument', 'prop.other'), names_to="target_type", values_to="proportion")

# Summarize the condition-level data for a barplot.
batch1z.first.move.summary <- batch1z.first.move.subject.summary %>%
  group_by(compatibility, target_type) %>%
  summarize(M=mean(proportion), SE=sd(proportion)/sqrt(n())) %>% 
  mutate(bias = factor(compatibility, levels = c("instrument", "equibiased", "modifier") ))
```

ggplot(batch1z.first.move.summary, aes(x=bias, fill=target_type, y=M, ymin=M-SE,ymax=M+SE))+
  geom_col(position=position_dodge(width=0.9), color = "black")+
  geom_errorbar(position=position_dodge(width=0.9), width=0.2)+
  scale_fill_brewer(palette = "Set1", name = "Location", labels = c("Animal", "Instrument", "Other"))+
  coord_cartesian(ylim = c(0,1))+
  theme_classic()+
  labs(y = "Proportion of first mouse movements", x = NULL)
```

for (r in 1:length(batch1$trial_index)){
  #print(r)
  #print(add_order(batch1$trial_index[r], batch1$trial_index[r-1], batch1$trial_order[r-1],verbose=T))
  batch1$trial_order[r] = add_order(batch1$trial_index[r], batch1$trial_index[r-1], batch1$trial_order[r-1],F)

}

# now connecting trials to each other based on elapsed time and rt

#View(batch1 %>% select(trial_index, internal_node_id,user_id, trial_order))
batch1a = batch1 %>%   
    arrange(user_id, trial_order, time_elapsed) %>% 
    mutate(prev_trial_end = time_elapsed - rt - 2) %>% 
    group_by(user_id, trial_index) %>% 
    mutate(daily_subjnum = 1:n(),
            unique_id = str_c(daily_subjnum,"__", user_id)) #%>% 
    # mutate(prev_trial_id = case_when(
    #   prev_trial_end == time_elapsed
    # ))
    # ungroup() %>% 
    # arrange(unique_id, trial_index)

link_trials  = function(s, to, df, tol=1, verbose=F){

  prev = df %>% filter(user_id == s & trial_order == to-1)
  curr = df %>% filter(user_id == s & trial_order == to)
  linked = vector("character", length(curr$prev_trial_end))

  for (q in 1:length(curr$prev_trial_end)){
    if(verbose){cat("q", q, ": ", curr$prev_trial_end[q],"\n")}
    for (p in 1:length(prev$time_elapsed)){
      if(verbose){cat("p", p, ": ", prev$time_elapsed[p],"\n")}
      if (!is.na(prev$time_elapsed[p]) & !is.na(curr$prev_trial_end[q])){
        if (curr$prev_trial_end[q]==prev$time_elapsed[p]| 
            (curr$prev_trial_end[q]<=prev$time_elapsed[p]+tol &
            curr$prev_trial_end[q]>=prev$time_elapsed[p]-tol)){
          if(verbose){cat("p",p, ": ", prev$unique_id[p],"\n")}
          linked[q] = prev$unique_id[p]
        }
      }
    }
  }
  return(linked)
}

#link_trials('6a935167-a0f5-4399-b6dd-8928fe8bae35',2,batch1a, tol=1)
link_trials('0f3402df-0da9-4a51-ac92-6a8c8f8e8c48',12,batch1a, tol=1)

linked_ids = vector("character")
for(user in unique(batch1a$user_id)){
  cat("\n\nUSER: ",user,"\n\n")
  for (trial in unique(batch1a[batch1a$user_id == user,]$trial_order)){
    cat("trial: ", trial,"\n")
    if(trial == 0){
      linked_ids = c(linked_ids, batch1a[batch1a$user_id == user & batch1a$trial_order == 0,]$unique_id)
    }else{
      linked_ids = c(linked_ids, link_trials(s=user, to = trial, df=batch1a, tol=1))
    }

  }
}

batch1b = batch1a
batch1b$linked_unique_ids  = linked_ids

dayx = batch1b %>% filter(user_id == 'c402ba0c-7a9d-4258-addd-149223724707')  %>% select(-starts_with('webgazer'))
```

#OKAY NOW NEED TO FILL IN LINKED IDS FOR THE CRITICAL TRIALS USING RTS FROM MOUSE CLICKS???

batch1.mouse = batch1b %>% 
  unnest(mouse_events,keep_empty = T) %>% 
  # mutate(prev_elapsed = time_elapsed - rt, 
  #          rt_eye = audio_start_time + t) %>% 
  group_by(user_id,trial_index,time_elapsed) %>% 
  mutate(prev_elapsed = time_elapsed-max(t))

dayx.mouse = batch1.mouse %>% filter(user_id == 'c402ba0c-7a9d-4258-addd-149223724707') %>% 
  select(-starts_with('webgazer')) %>% 
  select(rt, trial_index,trial_type, internal_node_id, time_elapsed, audio_start_time, type, object, t, prev_elapsed)
# 
# write_csv(dayx, "data/in-lab-merced/test.csv")

dayx.summ = dayx %>% group_by(internal_node_id) %>% summarize(n())

batch1.summary = batch1 %>% 
  dplyr::filter(compatibility != 'NA', compatibility != 'filler') %>%
  group_by(user_id, daily_subjnum, unique_id) %>% 
  summarize(n=n()) %>% 
  group_by(user_id) %>% 
  mutate(daily_total = max(daily_subjnum)) %>% 
  arrange( user_id, daily_subjnum)

```{r}
batch1.gaze = batch1b %>% 
  tidyr::unpack(webgazer_targets) %>%
  tidyr::unpack(c(`#jspsych-free-sort-draggable-0`, `#jspsych-free-sort-draggable-1`, `#jspsych-free-sort-draggable-2`, `#jspsych-free-sort-draggable-3`), names_sep=".") %>%
  unnest_wider(webgazer_data) %>% 
  unnest_longer(col = c(x, y, t))

dayx.gaze = batch1.gaze %>% filter(user_id == 'c402ba0c-7a9d-4258-addd-149223724707') %>% 
  ungroup() %>% 
  mutate(trial_lag = as.numeric(time_elapsed) - as.numeric(lag(time_elapsed))
        # ,last_gaze = t[[1]][length(t[[1]])]
         )

for (i in 1:length(dayx.gaze$t)){
  g = dayx.gaze$t[i][[1]][length(dayx.gaze$t[i][[1]])]
  if(is.null(g)){
    last_gaze[i]<-NA
  }else{
    last_gaze[i]<-g
  }
}

dayx.gaze = cbind(dayx.gaze,last_gaze) 
dayx.gaze = dayx.gaze %>%  
  mutate(extra = trial_lag - last_gaze) %>% 
  select(rt, trial_index, trial_type,time_elapsed,trial_lag,audio_start_time,t,last_gaze, extra)
```

#batch2

```{r E4-in-lab-load-data-second-batch}
# second batch

user.data.files <- list.files('data/in-lab-merced', pattern = "pushkin_users_*",full.names = TRUE)

all.user.data <- map_dfr(user.data.files, readRDS) %>% unique()

@@

#data.tables = readRDS("data/in-lab-merced/lex_stimulusResponses_2023-05-01.rds")
all.data <- map_dfr(data.files, readRDS) %>% 
  bind_rows(readRDS("data/in-lab-merced/verb_bias_ET_data_10_13_2022.rds")) %>% 
  arrange(user_id, trial_index, time_elapsed) %>% 
  group_by(user_id, trial_index) %>% 
  mutate(daily_subjnum = 1:n(),
         unique_id = str_c(daily_subjnum,"__", user_id)) %>% 
  ungroup() %>% 
  arrange(unique_id, trial_index) %>% 
  left_join(all.user.data, by = "user_id") %>% 
 # separate(created_at, into = c("date", "time"), sep = " ") %>% 
  mutate(date_time = parse_date_time(created_at, "ymd HMS")) %>% 
  mutate( date = date(date_time))

@@

*TO DO:* cross ref with angela's list of participant times/dates
-still trying to figure out how the daes/times line up...

participant_log = read_csv('data/in-lab-merced/participant_log.csv') %>% 
  separate(time, into = c("start_time", "end_time"), sep = "-") %>% 
  mutate(end_date_time = paste(date, end_time, sep = " "),
        PSTtime = strptime(end_date_time, format = "%m/%d/%y %H:%M", tz = "America/Los_Angeles"),
       # UTCtime = strptime(start_date_time, format = "%m/%d/%y %H:%M", tz = "America/New_York")
       other_tz_time = with_tz(PSTtime, "US/Alaska")
        ) %>% 
         #PSTtime = strptime(start_time, format = "%H:%M", tz = ""),
        # UTCtime = strptime(start_time, format = "%H:%M", tz = "GMT")) %>% 
  filter(action == "credit granted") %>% 
  group_by(date) %>% 
  arrange(PSTtime) %>% 
  mutate(order = 1,
         order = cumsum(order))

write_csv(participant_log, "data/in-lab-merced/participant_log2.csv")

all.user.data.summary = all.data %>% 
  dplyr::filter(compatibility != 'NA', compatibility != 'filler') %>%
  group_by(date_time, user_id, daily_subjnum, unique_id) %>% 
  summarize(n=n()) %>% 
  arrange(date_time,daily_subjnum)

write_csv(all.user.data.summary, "data/in-lab-merced/pushkin_user_data_summary.csv")
#hist(all.user.data.summary$n)

#all.user.data.summary %>% ungroup() %>% summarize(n_distinct(date_time), n_distinct(date))
```

### 2023-08-24 manuscript.Rmd
updated intro

Finally, @slim2022moving used the Webgazer.js library via the PCIbex online experiment platform [@zehr2018PCIbex] to replicate a visual world paradigm study in which participants (n=90) listened to sentences (e.g., Mary reads a letter)  while viewing four pictures displayed across the four quadrants of the screen (e.g., a letter, a backpack, a car and a wheelchair)  [@dijkgraaf2017predicting]. When the verb in the sentence was _constraining_ (e.g., read) with respect to the display (only the letter would be an appropriate continuation), listeners made more fixations to the target image (letter) than when the verb was _neutral_ (e.g., Mary steals a letter), as in previous work. However, they observed a substantial delay of ~200ms in the onset of the effect relative to the in-lab study with an infrared eyetracker and a reduction in effect size (despite a threefold increase in sample size relative to the original). The authors noted that delay was particularly problematic given that the purpose of the original study was to capture aspects of predictive or anticipatory processing: in the original study, but not in the web replication, the difference between _constraining_ and _neutral_ conditions emerged before the onset of the final noun (e.g., letter).

@@

In order to validate online eyetracking measures, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science. Ideally, we would only attempt to replicate studies where the original measurements have small error bars and are known to replicate; otherwise, it can be difficult to distinguish a failure of the method (online eyetracking does not work) from a failure of the original study to replicate.

In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several ``sanity check'' analyses, such as the correlation between calibration accuracy and effect size. (If the effect is real but there is noise from low-accuracy eyetracking, this correlation should be substantial.) Third, for XXX of the studies, we had comparison data collected in-lab either using jsPsych or a more traditional eyetracker technology, allowing us to direct assess the impact of differences in subject population and equipment.

@@

<!-- since there are not the analysis scripts we're using, I'm not sure about the best way to link to this -->
<!--AJ: I don't know what the above comment means. -->

All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/s82kz>.

>>>>

Finally, several papers have used the WebGazer.js library (via different interfaces) to replicate visual world paradigm studies. @slim2022moving used the PCIbex online experiment platform [@zehr2018PCIbex] to replicate a study in which participants (n=90) listened to sentences (e.g., Mary reads a letter)  while viewing four pictures displayed across the four quadrants of the screen (e.g., a letter, a backpack, a car and a wheelchair)  [@dijkgraaf2017predicting]. When the verb in the sentence was _constraining_ (e.g., read) with respect to the display (only the letter would be an appropriate continuation), listeners made more fixations to the target image (letter) than when the verb was _neutral_ (e.g., Mary steals a letter), as in previous work. However, they observed a substantial delay of ~200ms in the onset of the effect relative to the in-lab study with an infrared eyetracker and a reduction in effect size (despite a threefold increase in sample size relative to the original). The authors noted that delay was particularly problematic given that the purpose of the original study was to capture aspects of predictive or anticipatory processing: in the original study, but not in the web replication, the difference between _constraining_ and _neutral_ conditions emerged before the onset of the final noun (e.g., letter). Similarly, @degen2021seeing reported that effects of scalar implicature processing (comparing looks to the target in a four quadrant display for sentences such as "Click on the girl that has _some_ apples" vs. "Click on the girl that has _three_ apples" ) were smaller and delayed relative to those observed in the lab [@sun2020another].

In contrast, using the Gorilla experiment platform [@anwyl2020gorilla], @prystauka2023online observed robust effects of verb semantics and lexical interference in similar time windows to previous in-lab studies [@altmannIncrementalInterpretationVerbs1999;@kukona2014lexical], though a direct comparison was not possible because their studies used different materials than the in-lab experiments they were conceptually replicating. Furthermore, using jsPsych, @vos2022comparing closely replicated the magnitude and timecourse of effects in a lab-based visual world paradigm where the ROIs consisted of the left and right halves of the display, suggesting that the concerns about poor temporal resolution can be mitigated. Indeed, the eye-tracking plug-in in jsPsych uses a fork of Webgazer.js in which certain modifications have been made to minimize processing time. ^[See discussion at https://github.com/jspsych/jsPsych/discussions/1892]

In sum, webcam eye-tracking has been used to varying degrees of success across a small set of paradigms that are used in cognitive science research. One configuration, jsPsych with a modification of WebGazer.js, appears to have circumvented initial limitations in temporal precision [@krajbich2010visual; @vos2022comparing] but has only been tested with paradigms with minimal requirements in terms of spatial precision (both experiments used the two sides of the display as the ROIs). In the current work, we evaluate a broad variety of paradigms in terms of their suitability for webcam eye-tracking, in order to provide guidance for cognitive scientists aiming to increase and diversify participation in their eye-tracking-based research.

@@

In order to validate the online eyetracking methodology, with the particular configuration known to have the greatest temporal precision, jsPsych and a modification of Webgazer, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science. Ideally, we would only attempt to replicate studies where the original measurements have small standard errors and are known to replicate; otherwise, it can be difficult to distinguish a failure of the method (online eyetracking does not work) from a failure of the original study to replicate.

In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several ``sanity check'' analyses, such as the correlation between calibration accuracy and effect size. (If the effect is real but there is noise from low-accuracy eyetracking, this correlation should be substantial.) Third, for two of the studies, we had comparison data collected in-lab either using jsPsych or a more traditional eyetracker technology, allowing us to direct assess the impact of differences in subject population and equipment.

@@

<!-- since there are not the analysis scripts we're using, I'm not sure about the best way to link to this -->
<!--AJ: I don't know what the above comment means. -->
<!--RR: I think it's referencing that the scripts in the OSF repository are the ones the students use for their class projects and not the ones that produced the results in the paper which is maybe confusing but I can add some description above about how these data were collected in a classrooom setting. -->
All stimuli, experiment scripts, data, analysis scripts, and a
pre-registration are available on the Open Science Framework at
<https://osf.io/s82kz>.
